paper_replication

### 2021-12-15 paper_replication.Rmd
Added result tables integer restrictions, 50 state

```{r}
########### clean county dem data ###########
countydem <- read.csv("county_demographics.csv")

@@

```{r}
########### calculate abortion rates in 2010 ###########

@@

```{r}
########### calculate percent democratic votes by county in in 2008 ###########
countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)

@@

```{r}
########### calculate distances from county centroids to closest provider location ###########

@@

```{r}
#make new variable for high vs. low restrictions
restrictions <- restrictions %>% mutate(totalrestrict = parental_involvement + two_trips + trap + unconstitutional_ban, 
                                        restrictionlevel = case_when(totalrestrict >= 3 ~ "high", 
                                                              totalrestrict < 3 ~ "low"))
restrictions$state[!(restrictions$state %in% countydem$state)] #none
countydem <- left_join(countydem, restrictions, by = "state")
countydem$restrictionlevel <- factor(countydem$restrictionlevel, levels = c("low", "high"))

@@

```{r}
#look at demographics

@@

summarytable <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Abortion Data Included in Analyses by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

```{r}
#calculate propensity scores
propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
summary(propensitywts)

@@

# ggplot(df, aes(x = ps, fill = inanalysislabel, color = inanalysislabel)) +
#   geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
#   facet_grid(as.factor(inanalysis) ~ .) +
#   xlab('Probability of Being in the Analysis') +
#   ggtitle('Propensity Score Distribution by Group') +
#   scale_fill_discrete('') +
#   scale_color_discrete('') +
#   theme(legend.position = 'bottom', legend.direction = 'vertical')
# 
# df %>%
#   mutate(ps.grp = round(ps/0.05) * 0.05) %>%
#   group_by(inanalysis, ps.grp) %>%
#   summarize(n = n()) %>%
#   ungroup() %>%
#   mutate(n2 = ifelse(inanalysis == 0, yes = n, no =  -1*n)) %>%
#   ggplot(aes(x = ps.grp, y = n2, fill = as.factor(inanalysis))) +
#   geom_bar(stat = 'identity', position = 'identity') +
#   geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(inanalysis == 0, 8, -8))) +
#   xlab('Probability of Being in the Analysis') +
#   ylab('N') +
#   ggtitle('Propensity Score Distribution by Group') +
#   scale_fill_discrete(name = "", labels = c("Counties in States With Abortion Data", "Counties in States Without Abortion Data")) +
#   scale_x_continuous(breaks = seq(0, 1, 0.05)) +
#   theme(legend.position = 'bottom', legend.direction = 'vertical',
#         axis.ticks.y = element_blank(),
#         axis.text.y = element_blank())

@@

```{r}
#run models
countydemanalysis <- df %>% filter(inanalysis == 1)

@@

```{r}
####### add in natality, maternalmortality, and infantdeath data #######
natality <- read.delim("natalitydata.txt")
maternalmortality <- read.delim("maternalmortality.txt")
infantdeath <- read.delim("infantdeathrecords.txt")

@@

```{r}
#### recode restriction level and include all counties with abortion data in analysis ####

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Abortion Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

#run model
countydemanalysis <- df %>% filter(hasabortiondata == 1)

model <- lm(AbortionRate2010 ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
summary(model)

@@

```{r}
###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
hist(countydem$maternaldeathrate)

@@

```
```{r}
### repeat analysis for infant death rate ###
countydem <- countydem %>% mutate(hasinfantdeathdata = case_when(is.na(countydem$infantdeathrate) ~ 0, 
                                                              !(is.na(countydem$infantdeathrate)) ~ 1))

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(infantdeathrate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable %>% mutate(percent = infantdeathrate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

plot2 <- ggplot(df, aes(x = ps, fill = hasinfantdeathdatalabel, color = hasinfantdeathdatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasinfantdeathdata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')
print(plot2)

plot3 <- df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasinfantdeathdata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasinfantdeathdata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasinfantdeathdata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasinfantdeathdata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
print(plot3)

@@

```{r}
### repeat analysis for birth rate ###

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(Birth.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
usedcounties
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = Birth.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

plot2 <- ggplot(df, aes(x = ps, fill = hasbirthratedatalabel, color = hasbirthratedatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasbirthratedata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')
print(plot2)

plot3 <- df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasbirthratedata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasbirthratedata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasbirthratedata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasbirthratedata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
print(plot3)

@@

```{r}
### repeat analysis for fertility rate ###

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(Fertility.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
usedcounties
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = Fertility.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

plot2 <- ggplot(df, aes(x = ps, fill = hasfertilityratedatalabel, color = hasfertilityratedatalabel)) +
  geom_histogram(alpha = 0.3, position = 'identity', bins=15) +
  facet_grid(as.factor(hasfertilityratedata) ~ .) +
  xlab('Probability of Being in the Analysis') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete('') +
  scale_color_discrete('') +
  theme(legend.position = 'bottom', legend.direction = 'vertical')
print(plot2)

plot3 <- df %>%
  mutate(ps.grp = round(ps/0.05) * 0.05) %>%
  group_by(hasfertilityratedata, ps.grp) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(n2 = ifelse(hasfertilityratedata == 0, yes = n, no =  -1*n)) %>%
  ggplot(aes(x = ps.grp, y = n2, fill = as.factor(hasfertilityratedata))) +
  geom_bar(stat = 'identity', position = 'identity') +
  geom_text(aes(label = n, x = ps.grp, y = n2 + ifelse(hasfertilityratedata == 0, 8, -8))) +
  xlab('Probability of Being in the Analysis') +
  ylab('N') +
  ggtitle('Propensity Score Distribution by Group') +
  scale_fill_discrete(name = "", labels = c("Counties in States With Data", "Counties in States Without Data")) +
  scale_x_continuous(breaks = seq(0, 1, 0.05)) +
  theme(legend.position = 'bottom', legend.direction = 'vertical',
        axis.ticks.y = element_blank(),
        axis.text.y = element_blank())
print(plot3)

>>>>

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### clean county dem data ###########
countydem <- read.csv("county_demographics.csv")

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### calculate abortion rates in 2010 ###########

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### calculate percent democratic votes by county in in 2008 ###########
countypresdem2008 <- countypres %>% filter(party == "DEMOCRAT" & year == 2008)
countypresdem2008$county_name <- str_to_title(countypresdem2008$county_name)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
########### calculate distances from county centroids to closest provider location ###########

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#make new variable for high vs. low restrictions
restrictions <- restrictions %>% mutate(totalrestrict = parental_involvement + two_trips + trap + unconstitutional_ban, 
                                        restrictionlevel = case_when(totalrestrict >= 3 ~ "high", 
                                                              totalrestrict < 3 ~ "low"))
restrictions$state[!(restrictions$state %in% countydem$state)] #none
countydem <- left_join(countydem, restrictions, by = "state")
countydem$restrictionlevel <- factor(countydem$restrictionlevel, levels = c("low", "high"))

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#look at demographics

@@

summarytable <- countydemanalysis %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydemanalysis, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Abortion Data Included in Analyses by State") %>% kable_styling()
# names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#calculate propensity scores
propensitywts <- glm(inanalysis ~ Black.pct + Native.American.pct + Asian.pct + Hispanic.pct + Median.Income + Women + percentdem2008, data=countydem, family=binomial())
summary(propensitywts)

@@

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#run models
countydemanalysis <- df %>% filter(inanalysis == 1)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
####### add in natality, maternalmortality, and infantdeath data #######
natality <- read.delim("natalitydata.txt")
maternalmortality <- read.delim("maternalmortality.txt")
infantdeath <- read.delim("infantdeathrecords.txt")

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
#### recode restriction level and include all counties with abortion data in analysis ####

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(AbortionRate2010 ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = AbortionRate2010/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Abortion Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Abortion Data Included in Analysis by State") %>% kable_styling()
# names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

#run models, all 50 states, integer restrictions
countydemanalysis <- df %>% filter(hasabortiondata == 1)

ext_model_dist <- lm(AbortionRate2010 ~ totalrestrict + distancetoprovider + totalrestrict*distancetoprovider, data = countydemanalysis, weights = ps)
summ_ext_dist <- summary(ext_model_dist)

ext_model_no_dist <- lm(AbortionRate2010 ~ totalrestrict, data = countydemanalysis, weights = ps)
summ_ext_no_dist <- summary(ext_model_no_dist)

# Model Summaries (Note coefficients are per 1000 women)
table_ext_no_dist <- tibble(Characteristic = "Number of Types of Restrictive Legislation",
                  Beta = ext_model_no_dist$coefficients[2]*1000,
                  "p-value" = summ_ext_no_dist$coefficients[2,4]) %>% 
  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model without Distances, all 50 States",
          digits = 3)
print(table_ext_no_dist)

table_ext_dist <- tibble(Characteristic = c("Number of Types of Restrictive Legislation", 
                                            "Distance to Provider", 
                                            "Distance/Restriction Interaction"),
                  Beta = ext_model_dist$coefficients[2:4]*1000,
                  "p-value" = summ_ext_dist$coefficients[2:4,4]) %>% 
  kbl(caption = "Table ****: Change in Abortions per 1000 Women in Model with Distances, all 50 States", digits = 3)
print(table_ext_dist)

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
###further clean up maternaldeathrate, infantdeathrate, birthrate, and fertilityrate data ####
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "[()]", replacement = "")
countydem$maternaldeathrate <- gsub(countydem$maternaldeathrate, pattern = "Unreliable", replacement = "")
countydem$maternaldeathrate <- as.numeric(countydem$maternaldeathrate)
hist(countydem$maternaldeathrate)

@@

```
```{r, echo=FALSE, message=FALSE, warning=FALSE}
### repeat analysis for infant death rate ###
countydem <- countydem %>% mutate(hasinfantdeathdata = case_when(is.na(countydem$infantdeathrate) ~ 0, 
                                                              !(is.na(countydem$infantdeathrate)) ~ 1))

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(infantdeathrate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable %>% mutate(percent = infantdeathrate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
# names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
### repeat analysis for birth rate ###

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(Birth.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
usedcounties
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = Birth.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
# names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@

@@

```{r, echo=FALSE, message=FALSE, warning=FALSE}
### repeat analysis for fertility rate ###

@@

summarytable <- countydem %>% group_by(state) %>% summarize(totalcounties = n())
usedcounties <- aggregate(Fertility.Rate ~ state, data = countydem, function(x) {sum(!(is.na(x)))}, na.action = NULL)
usedcounties
summarytable <- left_join(summarytable, usedcounties, by = "state")
summarytable
summarytable %>% mutate(percent = Fertility.Rate/totalcounties) %>% kbl(col.names = c("State", "Total Counties", "Counties with Data in 2010", "Percent"), caption = "Table 1. Number of Counties with Data Included in Analysis by State") %>% kable_styling()
# names(countydem)
countydem <- countydem %>% mutate(Total = White + Black + Native.American + Asian + Hispanic, White.pct = White/Total, Black.pct = Black/Total, Native.American.pct = Native.American/Total, Asian.pct = Asian/Total, Hispanic.pct = Hispanic/Total, College.Grad.pct = College.Grad/Total)

@@
